---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# multnutFW

<!-- badges: start -->

<!-- badges: end -->

The goal of multnutFW is to analyze food web models that contain a user-defined number of essential chemical elements at fixed proportions of the biomass of each node.

## Installation

You can install the development version of multnutFW from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("robertwbuchkowski/multnutFW@development")
```

## Example

This is a basic example analyzing the introductory food web:

```{r example1}
library(multnutFW)

# Run the food web analysis of the introductory community:
intro_analysis = comana(intro_comm)

# The analysis is run on the food web included in the package with this feeding matrix:
intro_comm$imat

# The introductory community has 4 chemical elements:
names(intro_comm$prop)

```

This introductory analysis outputs several important features of the food web. First, it outputs the total consumption rate for each species in units of carbon.

```{r example2}
intro_analysis$consumption
```

It also outputs matrices of carbon (and any other nutrient) flow throughout the food web. To get the output matrix for another element, simply put that element's name in the place of "Carbon".

```{r example3}
intro_analysis$fmat$Carbon
```

It also outputs the mineralization rate for each species and each element. We can make a graph to show these data more clearly.

```{r example4}
# Data on mineralization rates:
intro_analysis$mineralization
```

Notice how some of the mineralization rates are negative. Microbe1 can immobilize nitrogen (canIMM = 1), but the other taxa cannot.

The package offers two options: correct diet or correct respiration.

To correct the diet, organisms shift their diet towards less abundant, nutrient rich food. Notice the changes in feeding preferences.

```{r example5}
# Correct the diet
intro_comm_diet = correct_diet(intro_comm)

intro_comm_diet$imat
```

The other option is to correct respiration by increasing the overflow respiration $\hat{E}_{C,i}$ term from 0 to whatever value is needed to remove the excess carbon. This change has to occur simultaneously with changes in feeding rate, because it often means that more volume needs to be eaten to overcome nutrient limitation.

Whenever the respiration term is reduced, this means that the organism is suffering elemental limitation from something other than carbon. The function automatically prints the element that is most limiting for each species.

```{r example6}
# Correct the respiration
intro_comm_resp = correct_respiration(intro_comm)

# Look at the new values for overflow respiration:
intro_comm_resp$prop$general$Carbon
```

### Calculating effects on nutrient mineralization

You can calculate the direct and indirect effects of each organism on mineralization using the built in function. This function allows you to customize the output to include whichever food web nodes and elements that are of interest.

```{r}
# Calculate the mineralization rates for all elements using the community with corrected respiration rates:
whomineralizes(intro_comm_resp)
```

### Using your own food web

You can run the analyzes on your own food webs easily by using the `build_foodweb` function to get them into the right format.

The function requires two inputs. First is a feeding list of predator-prey relationships. This data frame also contains the rates of assimilation efficiency for each trophic interaction coded with the letter 'a' before each element included in the model. Sometimes, users would like to use the same assimilation efficiency for each food source or for each node. You can do this by just repeating the assimilation efficiency for each feeding interaction. One easy way to implement this is to join a table of node ID and assimilation efficiency with the feeding list table by either Predator or Prey columns, respectively.

```{r example7}
# Feeding list example for the introductory community:
feedinglist
```

The second is a properties dataframe listing the necessary parameters.

```{r example8}
# Feeding list example for the introductory community:
head(properties_example1)
# Only printing the head of this file as an example.
```

These two files can be put into the function to create your new food web.

```{r example9}
yourfoodweb = build_foodweb(feeding = feedinglist, properties = properties_example1)

str(yourfoodweb)
```

The function creates an interaction matrix and also breaks out the assimilation efficiency and properties by element into the appropriate matrices for the package. If you look at the structure of this community, you can see how to build it manually using lists.

### Modifying respiration parameters

The model contains respiration parameters that either take respiration as a function of biomass $E$ or as a function of consumption rate $p$. The package contains functions to switch between these two parameters during the calculations in $comana$. At equilibrium, the difference does not matter, but it is important when simulating the food web. Simulations will be explored next.

```{r example10}

# Take the introdutory community and switch respiration by E to p.
temp_comm = E_to_p(intro_comm)

temp_comm$prop$general$Carbon

# Switch back
temp_comm = p_to_E(temp_comm)

temp_comm$prop$general$Carbon

# Clean up environment
rm(temp_comm)
```
